---
title: "Data Visualization"
author: "Javier Patron"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggbeeswarm)
library(gapminder)
library(sf)
```

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

```{r}
ggplot(data = lizards, aes(x=total_length, y = weight))+
  geom_point()
```

Create a histogram for total length

```{r}
plot1 <- ggplot(lizards, aes(x = total_length)) + 
  geom_histogram(aes(color = site,
                     size = 1,
                     linetype= "--"))

plot1

```

```{r}

ggplot(data = lizards, aes(x=total_length, y = weight))+
  geom_point(shape = 22)

```

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = common_name, 
                size = total_length),
             alpha = 0.5)
```

```{r}

ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(color = "darkblue") +
  facet_wrap(~common_name)
```

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point() +
  facet_grid(~common_name)

```

total amount of lizard by just using common name

```{r}
table <- lizards |> 
  group_by(common_name) |> 
  summarise(count = n())

```

graph your new table

```{r}
ggplot(table, aes(y = fct_reorder(common_name, count),
                  x = count)) +
  geom_col()
```

```{r}

ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_wrap(~ scientific_name) + 
  theme_dark() + 
  labs(title = "Cool Lizards",
       subtitle = "Total Length vs. Weight",
       x = "Total Length (mm)",
       y = "Weight (g)")


```

```{r}
 x <-list(numbers = c(1,2,3), letters = c("a", "b"))

unlist(x, recursive = FALSE, use.names = FALSE)
```

```{r}



```

## Scales

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high = "red")

                        

```

Create your own pallett of gradient colors

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue", "green", "yellow"))
```

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "orange", 
                    mid = "red", 
                    high = "black",
                    midpoint = 150,
                    breaks = c(50,75,150))
```

```{r}
ggplot(data = lizards, aes(x=total_length, 
                           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, 
                                  to = 300,
                                  by = 60))
```

Make a horizontal boxplot with common name on the y-axis, total length on the x-axis, with boxplot fill changing based on the common name

new column, rename the column or renaming, order the names,

```{r}

lizards_median <- lizards |> 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))


ggplot(lizards, aes(y=common_name, x= total_length)) +
  geom_boxplot(aes(fill = common_name))


ggplot(lizards_median, aes(y= common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "beyonce::X1") +
  theme_minimal()

```

## Make things look bad

```{r}
ggplot(data = lizards, aes(x=total_length, y = weight)) + 
  geom_point(aes(color = weight))+
  theme_dark()+
  theme(panel.grid = element_line(color = "white"),
        panel.grid.minor.y = element_blank(),
        axis.title.x = element_text(color = "magenta"),
        axis.text.y = element_text(color = "green"),
        text = element_text (size = 20),
        panel.background = element_rect(color = "red", fill = "white")) + 
  annotate("text", x = 300, y = 50, label = "whooooo", color = "blue") +
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)
```

Repulsive labels with `ggrepel`\`

```{r}

wws_lizards <- lizards |> 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = wws_lizards, aes(x=total_length, y = weight)) + 
  geom_point() + 
    geom_text_repel(aes(label = toe_num))
```

2002 and the continent is europe

```{r}
gapminder |> 
  filter(year == 2002, continent == "Europe") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)

ggplot(data = lizards, aes(x = total_length, y = weight )) +
  geom_point() +
  gghighlight(weight > 30, label_key = toe_num)
```

```{r}

jornada_veg <- read_sf(here::here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot(jornada_veg) +
  geom_sf(aes(fill = Name)) +
  theme_dark() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
  

```
